### 2015-05-13 whoop.tex
updates to optimisations

\ADComment{Briefly talk about how we tried inlining and do summarization instead}

@@

To generate the candidate invariants, \whoop performs an inter-procedural static analysis on the call-graph (in Boogie IVL code) of each entry point, identifying all possible accesses to shared memory locations. To be sound, if \whoop cannot infer information for a specific access, it generates a coarse-grained candidate for the corresponding memory region. For example the previous \texttt{\$M.0} candidates would now be:

\begin{boogie}
procedure foo$ep1(dev: int);
  ensures _b$ls$entrypoint1$0
    ==> WATCHED_ACCESS_$M.0 == dev + 4
    ==> lock$0_in_LS_$M.0_$entrypoint1;
  ensures _b$ls$entrypoint1$1
    ==> WATCHED_ACCESS_$M.0 == dev + 8
    ==> lock$0_in_LS_$M.0_$entrypoint1;
  ensures _b$ac$entrypoint1$0
    ==> WATCHED_ACCESS_$M.0 == dev + 4
    ==> !WRITTEN_$M.0_$entrypoint1;
  ensures _b$ac$entrypoint1$1
    ==> WATCHED_ACCESS_$M.0 == dev + 8
    ==> !WRITTEN_$M.0_$entrypoint1;
  ...
\end{boogie}

@@

\medskip\noindent\textbf{Optimizations }
%
To increase precision we enhance \whoop with domain-specific knowledge of how the Linux kernel interacts with its drivers. This is an ongoing manual effort: the more drivers we study, the more domain-specific properties we discover, which we then exploit to make \whoop more precise.

In our experience, one of the most effective precision optimizations is to enrich \whoop with information regarding kernel-imposed serialization. The Linux kernel can serialize calls to entry points, thus forcing them to not run concurrently with each other. An an example, a large number of networking and ethernet entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop exploits this knowledge and does not create pairs for entry points that are serialized.

Modeling kernel methods is another opportunity to increase precision, but requires Linux kernel expertise. For example, the \texttt{register\_netdev()} function registers a network driver with the kernel. Before this function is called, the kernel is not able to invoke any of the driver's entry points. Modeling this functionality can reduce false positives when race-checking accesses before the driver's registration.

Another optimization \whoop uses is to soundly reduce the checked memory locations. If a memory location is accessed only by one entry point in a pair, then it is safe to not check this location for this pair. This reduces the burden to the SMT solver and speed up the verification process.

>>>>

@@

To generate the candidate invariants, \whoop performs an inter-procedural static analysis on the call-graph (in Boogie IVL code) of each entry point, identifying all possible accesses to shared memory locations. To be sound, if \whoop cannot infer information for a specific access, it generates a coarse-grained candidate for the corresponding memory region.

@@

\medskip\noindent\textbf{Optimizations }
%
To increase precision we enhance \whoop with domain-specific knowledge of how the Linux kernel interacts with its drivers. This is an ongoing manual effort: the more drivers we study, the more domain-specific properties we discover, which we then exploit to make \whoop more precise. In our experience, one of the most effective precision optimizations is to enrich \whoop with information regarding kernel-imposed serialization. The Linux kernel can serialize calls to entry points, thus forcing them to run sequentially with each other. As an example, a large number of networking entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop exploits this knowledge and does not create pairs for entry points that are serialized.

%Modeling kernel methods is another opportunity to increase precision, but requires Linux kernel expertise. For example, the \texttt{register\_netdev()} function registers a network driver with the kernel. Before this function is called, the kernel is not able to invoke any of the driver's entry points. Modeling this functionality can reduce false positives when race-checking accesses before the driver's registration.

A performance optimization that \whoop uses is to soundly reduce the checked shared variables. If a shared variable $s$ is accessed only by one entry point in a pair, then it is safe to not check $s$ for this pair. This reduces the burden of the SMT solver and speeds up the verification process.
